Verification of Stencil Computations with Spatial Specifications

### 2016-11-08 paper.tex
rebuilding section on model

% SYNTAX OPERATIONS AND PREDICATES
\newcommand{\neigh}{\textsf{neigh}}
\newcommand{\arrayTy}{\textsf{array}}
\newcommand{\rhsExp}{\textsf{rhsExp}}
\newcommand{\var}{\textsf{var}}

@@

@@

>>>>

% SYNTAX OPERATIONS AND PREDICATES
%\newcommand{\neigh}{\textsf{neigh}}
\newcommand{\arrayTy}{\textsf{array}}
\newcommand{\rhsExp}{\textsf{rhsExp}}
\newcommand{\var}{\textsf{var}}

@@

\begin{definition}[Base induction variable]
  An integer variable is a \emph{base induction variable} if is the
  control variable of a ``for'' loop (or equivalent, \eg{},
  \texttt{do} in Fortran for our implementation) incremented by $1$
  per iteration. The variable is interpreted as a based induction
variable only within the scope of the loop body. For brevity we
say \emph{induction variable}, which are ranged over by $i, j, k$
throughout.
%TODO: consider mentioning derived, non-based induction variables
% as we did before.
\end{definition}

\begin{definition}[Array subscripts and indices]
  An \emph{array subscript}, denoted $a(\bar{e})$, is an expression
  which reads from an array $a$ at an \emph{index} specified by a
  comma-separated sequence of integer expressions denoted $\bar{e}$ or 
  in expanded form as $(e_1, \ldots, e_n)$. An
  \emph{index} $\bar{e}$ is called \emph{relative} if it is defined in
  terms of induction variables.
\end{definition}

\begin{definition}[Neighbourhood index]
\label{def:neighbour}
  For an array subscript $a(\bar{e})$ an index $\bar{e}$ is called a
  \emph{neighbourhood index} if every $e \in \bar{e}$ is of the form
  $e \equiv i$ or $e \equiv i + c$ or $e \equiv i - c$ where $c$ is an
  integer constant. That is, a neighbourhood index is a relative index
  where each component is a constant translation of a base induction
  variable. (The relation $\equiv$ here identifies terms up-to
  commutativity of $+$ and the inverse relationship of $+$ and $-$
  \eg{}, $(-b) + i \equiv i - b$).  %We classify neighbourhood indices
%  by the predicate $\neigh_I$ where $I$ is a set of induction
%  variables. We omit $I$ when clear from the context.
\end{definition}

@@

\begin{definition}
  An \emph{index scheme} is a vector of size $n$ (an $n$-vector) with
  values drawn from $\domainVal = \mathbb{Z} \cup \{\infty\}$.  Values
  in $\mathbb{Z}$ represent constant offsets of neighbourhood indices
  (expressions of the form $i \pm c$ where $i$ is an induction
  variable and $c$ is a constant integer, Def.~\ref{def:neighbour},
  p.~\pageref{def:neighbour}). $\infty$ represents any indexing
  behaviour--- it is a wildcard.  Throughout $u, v, w$ range over
  index schemes and $u_i$ denotes the $i^{th}$ element of $u$. We
  write $(\domainVal)^n$ for the set of $n$-dimensional schemes.
\end{definition}

Indexing schemes give an abstract
representation of array subscripts in the source language.
Our model represents the meaning of stencil specifications
as indexing schemes, which explain the range of possible
indexing behaviours which are consistent with the specification.
The relationship between indexing schemes and source language
terms is given precisely by the following function \textsf{schematic} which
maps syntactic index expressions to indexing schemes. This
is used in inference and checking.

\newcommand{\schemeOne}{\textsf{s}_I}
%\newcommand{\schemeOne}{\textsf{schema$^1_I$}}
\newcommand{\schemeN}{\textsf{schematic}_I}
%\newcommand{\schemeN}{\textsf{schema$^n_I$}}
\begin{definition}%[Schematic]
\label{def:schematic}
Let $\schemeOne$ be a partial function, parameterised by a set of
induction variables $I$, mapping expression syntax to
$\domainVal$ values:
%%
\begin{align*}
\schemeOne(e)
 & =  \begin{cases}
c & e \equiv i + c \;\; \wedge \;\; i \in I \\
-c & e \equiv i - c \;\; \wedge \;\; i \in I \\
\infty & \textit{$I \not\subseteq \textit{FV}(e)$} \\
\bot   & \textit{otherwise}
\end{cases}
\end{align*}
where $\textit{FV}(e)$ is the set of free variables in $e$. In the
first two cases, expressions that are a constant offset of an 
induction variable (\ie{}, neighbourhood indices) are mapped to the
offset $c$ or $-c$ $\in \domainVal$. In the third case, if the
expression is not a neighbourhood index, but neither is it a relative
index (\ie{}, it is not defined in terms of an induction variable) then it
is mapped to $\infty \in \domainVal{}$. Otherwise, $\schemeOne(e)$
is undefined (returns $\bot$).

The partial function $\schemeN$, also parameterised by a set of
induction variables $I$, uses the above operation $\schemeOne$ to
 maps from the syntax of indices to index schemes, \ie{} values of 
$\domainVal^n$. It is defined on the syntax of index
expressions $\bar{e} = (e_1, ..., e_n)$ as:
\begin{align*}
\schemeN(e_1, ..., e_n) & =
[\schemeOne(e_1) ... \schemeOne(e_n) ]
\quad \textit{iff} \;\; \forall x . \schemeOne(e_x) \neq \bot
\end{align*}
%
Thus, for an indexing expression, if each component is a neighbourhood
index, then $\schemeN$ maps it to an indexing scheme.
\end{definition}
